Intersection of non-matching grids of different dimensions

### 2016-09-26 paper_intersections.tex
New tracing table algorithm.
- can deal with all cases, but dealing with triangle
touching an edge is ugly

\subsection{Intersection Triangle-Tetrahedron (2D-3D)}
The intersection of a triangle $S_3$ and a tetrahedron $S_4$ is an $n$-side polygon, $n\le 7$. The sides of the polygon
are coincident with either sides of $S_3$ or with faces of the $S_4$. Thus each vertex of the polygon
can arise either from side-face intersection, or from edge-triangle intersection, or be a vertex of $S_3$.
So we have to compute at most 12 side-face intersections and at most 6 edge-triangle intersections. However,
to this end we only need to compute 9 \plucker coordinates (3 sides, 6 edges) and 18 permuted inner produts, one for every side-edge pair.
Computation of the intersection polygon consists of three parts: calculation of IPs on the $S_3$ boundary (Algorithm \ref{algo:colect_23_ip_triangle}),
calculation of IPs in the interior of $S_3$ (Algorithm \ref{algo:collect_23_ip_edges}), and finalize the correct order of IPs (Algorithm \ref{algo:order_23}).

@@

  \DontPrintSemicolon
  \KwIn{input data}
  \KwOut{List of IPs on sorted output data}  
  $F(:,:)=-1$\tcp*{Unset entries of the face connection table.}

  \For{side $i$ of $S_3$}{ 
    $L$ = intersection($i, S_4$)\;
    \For{ $p \in L$, $p \ne I[-1]$ }{append $p$ to $I$}    
    \lIf{$\abs{L}=2$}{
        append $0,0$ to $s$
    } \lElse {
        append $1$ to $s$
    }
  }
  \nllabel{line:ip_first_vtx_duplicate}
  \lIf{ $I[0] == I[-1]$ }{remove $I[-1]$}

  $l=0$
  \For{$p \in I$}{
    set $j$ to index of $p$ in $I$\;
    \If{ $p$ is on boundary of $S_4$}{
      $x$ is vertex,edge, or face on which $p$ lies\; 
      \For{ $i=0,\dots, s[j]$}{
        $F[l, x]=j$\;
        $l=1-l$ \tcp*{switch from input to output and vice versa}
      }  
    }
  }  
\end{algorithm}

In Algorithm \ref{algo:colect_23_ip_triangle}, we first compute IPs
on the boundary of $S_3$ using the line-tetrahedron intersection algorithm for every side. Every IP that is not same as the previous IP 
(occures at vertices of $S_3$) is appended to $I$. The IPs are added in the order around $S_3$ which match the orientation of the intersection polygon.
The array $s$ indicates IPs that comes from side touching $S_4$ at its edge or vertex. 
At line \ref{line:ip_first_vtx_duplicate}, we remove posible dupicate IP for the first vertex.
Then we pass through $I$ and denote to the table $F$ at which IP ($j$) the polygon's boundary switchs from the $S_4$ boundary to the $S_3$ boundary
($l=0$) and vice versa ($l=1$). The table $F$ contains the entry ($l=1$) and leave ($l=0$) IP for every geometrical object (vertex, edge, or  face) 
$x$ on the boundary. In particular the touching IPs appear in both directions. The result of Algorithm \ref{algo:colect_23_ip_triangle} is incomplete 
list $I$ of IPs lying on the $S_3$ boundary and the table $F$ stores relation of the objects on the $S_4$ boundary to the IPs on the $S_3$ boundary.

@@

  \nllabel{line:12edges}
  \lFor{edge $e$ of $S_4$}{$L[e]$ = intersection($e, S_3$)} 
  \For{unmarked edge $e$ of $S_4$}{
    \uIf{ $L[e]$ is inside of $e$}{ 
        \nllabel{line:edge_faces}        
        $(f_0,\ f_1)$ = edge faces($e$)
    } 
    \Else($p$ on the vertex $v$ of $S_4$){       
            $(f_0,\ f_1)$ = vertex faces($v$)   \tcp*{Algorithm \ref{algo:vertex_faces}}
    }    
    \uIf{ $L[e]$ is on boundary of $S_3$ }{
         \uIf{ $L[e]$ is inside of $e$}{ 
             get $i$ such that $F[i,e] = k$, $k\ne -1$
         }
         \Else($p$ on the vertex $v$ of $S_4$){ 
             get $i$ such that $F[i,v] = k$, $k\ne -1$          
         }         
         $F[1-i, f_{1-i}] = k$\;
    }
    \Else{
         append $L[e]$ to $I$\;
         set $j$ to index of $L[e]$ in $I$\;
         F[0, $f_0$]=j\;
         F[1, $f_1$]=j\;
         mark all edges coincident with $L[e]$
    }   
  }  
\end{algorithm}

@@

  \SetKw{return}{return}
  \DontPrintSemicolon
  \KwIn{vertex $v$ of $S_4$, $L[:]$ intersection results for edges of $S_4$}
  \KwOut{faces $S_4$ intersected by the plane of $S_3$}
  $e_0,\ e_1,\ e_2$ edges coincident with $v$\;
  $s[i] = L[e_i]$, for $i=0,1,2$\;
  \nllabel{line:two_degenerated}
  \uIf{$s$ have 1 non-degenerate edge $e$}{
    \return pair of degenerate edges sorted according to \edgefaces($e$)    
  }
  \nllabel{line:one_degenerated}
  \uElseIf{ $s$ have 1 degenerate edge $e$}{
    $f$ is face opposite to $e$first of faces coincident with $e$\;
    $g$ is non-degenerate edge coincident with $f$\;
    \return edge faces($g$)
  }
  \nllabel{line:different_sign}
  \uElseIf{ $s$ have edge $e$ with sign oposite to other two}{
    \return edge faces($e$)\;
  }
  \nllabel{line:same_signs}
  \Else($s$ have all signs same){
    \return anything\;
  }  
\end{algorithm}

@@

\subsubsection{Vertex Faces Algorithm}
This function get an IP $p$ at vertex $v$ of $S_4$ as a parameter. The IP is special vertex case of non-degenerate edge-triangle intersection. 
The functin returns pair of generalized faces of $S_4$ preceeding and succeeding $p$ on the polygons boundary in the case that $p$ is at interior of $S_3$.
Basic idea is to use the signs of IPs of the three edges coincident with $v$. 
Possible cases are:
\begin{itemize}
 \item {\bf All IPs have same sign.} (line \ref{line:same_signs}) We return any pair of faces.  $S_3$ is touching $S_4$ 
 in the vertex $v$, the polygon ddegenerates to the single IP $p$, no connectionn information from table $F$ necessary.
 \item {\bf Single IP have sign oposite to the other two.} (line \ref{line:different_sign}) Let $e$ be the edge of the single $IP$ with different sign. 
 The plane of $S_3$ separates $e$ from the other two edges so it goes through the faces adjacent to $e$. The order is detemined by the function edge faces.
 \item {\bf Single degenerated IP.} (line \ref{line:one_degenerated}) Let us denote $e$ the edge with degenerated IP and $f$ the face between the other two edges.
 The other two (non-degenerate) edges may have either oposite sign (the plane cut $S_4$) or the same sign (the plane touching $S_4$ at the edge $e$.

>>>>

\subsection{Intersection Triangle-Tetrahedron (2D-3D)}
The intersection of a triangle $S_3$ and a tetrahedron $S_4$ is an $n$-side polygon, $n\le 7$. The sides of the polygon
are coincident with either sides of $S_3$ or with faces of the $S_4$. Thus each vertex of the polygon
can arise either from side-face intersection, or from edge-triangle intersection, or be a vertex of $S_3$.
So we have to compute at most 12 side-face intersections and at most 6 edge-triangle intersections. However,
to this end we only need to compute 9 \plucker coordinates (3 sides, 6 edges) and 18 permuted inner produts, one for every side-edge pair.
Computation of the intersection polygon consists of three parts: calculation of IPs on the $S_3$ boundary (Algorithm \ref{algo:colect_23_ip_triangle}),
calculation of IPs in the interior of $S_3$ (Algorithm \ref{algo:collect_23_ip_edges}), and finalize the correct order of IPs (Algorithm \ref{algo:order_23}).
Correct order of the IPs stored in the list $I$ is defined by tables $F_g(:)$ and $F_p(:)$. Every side of the intersection polygon lies either on side of $S_3$, or
on face, or on edge of $S_4$. Let us denote $M_3$ the set of sides of $S_3$ and $M_4$ the set of volume, faces, edges, and vertices of $S_4$. 
Every side of polygon that lies on $x\in M_32$ is folowed by an IP given by $F_g[x]$ and every IP $p$ is folowed by the side that lies on $F_p[p]\in M_32$.

@@

  \DontPrintSemicolon
  \SetKw{continue}{continue}
  \KwIn{input data}
  \KwOut{List of IPs on sorted output data}  
  $F_g(:)=-1$\tcp*{Unset IPs of geometric entities.} 
  $F_p(:)=-1$\tcp*{Unset geometric entities of IPs.}

  \For{side $i$ of $S_3$}{ 
    $L$ = intersection($i, S_4$)\;
    \lIf{ \abs{L} = 0}{\continue}
    \lIf{ \abs{L} = 1}{
        append $p$ to $J$
        \continue
    }
    \For{$p$ in $L$}{
        \lIf{$p \ne I[-1]$ }{append $p$ to $I$}
        $p$ lies on $x\in M_4$ and side $s\in M_3$\; 
        \uIf{$p$ is first in $L$}{
            $F_g[x] = p$\;
            $F_p[p] = s$\;
        }
        \Else($p$ is the last in $L$){
            $F_g[s] = p$\;
            $F_p[p] = x$\;
            \nllabel{line:back_link43}
            \lIf{$F_g[x] = -1$}{$F_g[x] = p$}\;
        }
    }
  }  
  \lIf{$I[-1] = I[0]$}{
    $F_g[F_p[I[-1]]]]=I[0]$\;
    $F_g[s_2] = I[0]$\;
    remove $I[-1]$
  }
\end{algorithm}

In Algorithm \ref{algo:colect_23_ip_triangle}, we first compute IPs
on the boundary of $S_3$ using the line-tetrahedron intersection algorithm for every side. We store intersections with just single IP ito separate list $J$
and skip filling the connection tables. This happen when the side just touch $S_4$ in edge or vertex. 
These IPs will be rediscovered again in Algorithm \ref{algo:collect_23_ip_edges} with better topological information, however
this is not the case if the touched edge $e$ of $S_4$ is coplanar with $S_3$ and the IP is inside of $e$. 
Therefore we keep a separate list of IPs to deal with this case.
For the intersection $L$ we pass through its IPs. The IP $p$ is added to the list $I$ unless the last point is the same. 
This condition is effective just for the first point of $L$ and merge IPs at the same vertex of $S_3$. If the vertex $v$ is between sides $s_1$, $s_2$, 
we effectively set $F_g[s_1] = v$, $F_g[y]=v$, $F_p[v] = s_2$. The backward temporary link on the line \ref{line:back_link43} 
is used in Algorithm \ref{algo:collect_23_ip_edges} to fix $F_p[p]$ for the second $p$ lying on an edge or a vertex of $S_4$. The condition deals with the case
when two sides intersect the same object $x\in M_4$.

@@

  \nllabel{line:12edges}
  \lFor{edge $e$ of $S_4$}{$L[e]$ = intersection($e, S_3$)} 
  \For{unmarked edge $e$ of $S_4$}{
    $p=L[e]$\;
    \uIf{ $p$ is inside of $e$}{ 
        \nllabel{line:edge_faces}        
        $(f_0,\ f_1)$ = edge faces($e$)
    } 
    \Else($p$ on the vertex $v$ of $S_4$){       
            $(f_0,\ f_1)$ = vertex faces($v$)   \tcp*{Algorithm \ref{algo:vertex_faces}}
    }    
    \uIf{ $p$ is on boundary of $S_3$ }{
         $p\in x \in M_4$; $x$ is edge or vertex 
         $q=F_g[x]$\tcp*{$q$ is already computed $p$}
         \lIf{$F_p[q]=x$}{$F_p[q]=f_1$}
         \lElse{$F_g[f_0]=q$}
         $F_g[x]=-1$\tcp*{remove the backlink}
    }
    \Else{
         append $p$ to $I$\;
         $F_g[f_0]=p$\tcp*{overwrites the backlink}
         $F_p[p]=f_1$
         mark all edges coincident with $p$
    }   
  }

  \lIf{$\abs{I} < 3$}{ \return $J$}
  \lElse{ return $I$ soter according connectivity in $F_g$ and $F_p$}
\end{algorithm}

@@

  \SetKw{return}{return}
  \SetKw{edgefaces}{edge faces}
  \DontPrintSemicolon
  \KwIn{vertex $v$ of $S_4$, $L[:]$ intersection results for edges of $S_4$}
  \KwOut{$(x_1, x_2)$, $x_1, x_2 \in M_4$, coincident with $v$ and intersected by the plane of $S_3$}
  $e_0,\ e_1,\ e_2$ edges coincident with $v$ oriented out of $v$
  $s[i] = L[e_i]$, for $i=0,1,2$, \;
  \nllabel{line:two_degenerated}
  \uIf{$s[:]$ have 1 non-degenerate edge $e$}{
    \return pair of degenerate edges sorted according to \edgefaces($e$)    
  }
  \nllabel{line:one_degenerated}
  \uElseIf{ $s$ have 1 degenerate edge $e$}{
    $f$ is face oposite to $e$
    \uIf{other two edges $e_a$, $e_b$ have different sign}{
        $z=\edgefaces(e_a)$\;
        replace $g\in z$, $g\ne f$ with $e$
        \return $z$
    }\lElse{
        append IP of $v$ to $J$
        \return anything\;
    }
  }
  \nllabel{line:different_sign}
  \uElseIf{ $s$ have edge $e$ with sign oposite to other two}{
    \return edge faces($e$)\;
  }
  \nllabel{line:same_signs}
  \Else($s$ have all signs same){
    append IP of $v$ to $J$
    \return anything\;
  }  
\end{algorithm}

@@

\subsubsection{Vertex Faces Algorithm}
This function get an IP $p$ at vertex $v$ of $S_4$ as a parameter. The IP is special vertex case of non-degenerate edge-triangle intersection. 
The functin returns pair of generalized faces of $S_4$ preceeding and succeeding $p$ on the polygons boundary in the case that $p$ is at interior of $S_3$.
Basic idea is to use the signs of IPs of the three edges coincident with $v$. 
Possible cases are:
\begin{itemize}
 \item {\bf All IPs have same sign.} (line \ref{line:same_signs}) We return any pair of faces.  $S_3$ is touching $S_4$ 
 in the vertex $v$, the polygon ddegenerates to the single IP $p$, no connectionn information from table $F$ necessary.
 \item {\bf Single IP have sign oposite to the other two.} (line \ref{line:different_sign}) Let $e$ be the edge of the single $IP$ with different sign. 
 The plane of $S_3$ separates $e$ from the other two edges so it goes through the faces adjacent to $e$. The order is detemined by the function edge faces.
 \item {\bf Single degenerated IP.} (line \ref{line:one_degenerated}) Let us denote $e$ the edge with degenerated IP and $f$ the face between the other two edges.
 The other two (non-degenerate) edges may have either oposite sign (the plane cut $S_4$) or the same sign (the plane touching $S_4$ at the edge $e$).
 In the first case, the call of edge faces for $e$ returns $(f_x,f)$ or $(f,f_x)$, then thr vertex faces function returns $(e, f)$ or $(f, e)$, respectively.
